---
title: "Assignment 3 - Sentiment Analysis II"
author: "Marie Rivers"
date: "4/26/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, results='hide', message=FALSE, warning=FALSE}
library(quanteda)
library(quanteda.sentiment)
library(quanteda.textstats)
library(tidyverse)
library(tidytext)
library(lubridate)
library(wordcloud) #visualization of common words in the data set
library(reshape2)
library(sentimentr)
library(kableExtra)
```

This assignment uses tweet data for the term 'IPCC'
```{r tweet_data}
raw_tweets <- read.csv("https://raw.githubusercontent.com/MaRo406/EDS_231-text-sentiment/main/dat/IPCC_tweets_April1-10_sample.csv", header=TRUE)

dat<- raw_tweets[,c(4,6)] # Extract Date and Title fields

tweets <- tibble(text = dat$Title,
                  id = seq(1:length(dat$Title)),
                 date = as.Date(dat$Date,'%m/%d/%y'))

#clean up the URLs from the tweets (people linking to news articles and such)
tweets$text <- gsub("http[^[:space:]]*", "",tweets$text) # substitute http links with nothing
tweets$text <- str_to_lower(tweets$text)
```

# 1.  Think about how to further clean a twitter data set. Let's assume that the mentions of twitter accounts is not useful to us. Remove them from the text field of the tweets tibble.
```{r}
tweets_clean <- tweets %>% 
  mutate(text_clean = text)  # keeping a column of the original text as a check
  
tweets_clean$text_clean <- gsub("@[^[:space:]]*", "", tweets_clean$text_clean)
head(tweets_clean)
```

# 2.  Compare the ten most common terms in the tweets per day.  Do you notice anything interesting?
```{r}
#tokenize tweets to individual words
words <- tweets_clean %>%
  select(id, date, text_clean) %>%
  unnest_tokens(output = word, input = text_clean, token = "words") %>% 
  anti_join(stop_words, by = "word")
```

```{r}
words_count <- words %>% 
  count(date, word)

top_ten_per_day <- words_count %>% 
  group_by(date) %>% 
  top_n(10, n)
```

```{r}
top_ten_table = aggregate(top_ten_per_day$word, list(top_ten_per_day$date), paste, collapse=", ") %>% 
  rename(Date = Group.1) %>% 
  rename(top_words = x) %>% 
  kable(col.names = c("Date", "Top 10 Words")) %>% 
  kable_paper(full_width = TRUE) %>% 
  row_spec(c(0), background = "lightgray")
top_ten_table
```

# 3.  Adjust the wordcloud in the "wordcloud" chunk by coloring the positive and negative words so they are identifiable.
```{r warning=FALSE}
#load sentiment lexicon
bing_sent <- get_sentiments('bing')

words_sent <- words %>% 
  left_join(bing_sent, by = "word") %>%
  left_join(
    tribble(
      ~sentiment, ~sent_score,
      "positive", 1,
      "negative", -1),
    by = "sentiment")

wordcloud_sent <- words_sent %>% 
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("lightsalmon4", "dodgerblue3", "gray40"),
                   max.words = 100)
```

# 4. Let's say we are interested in the most prominent entities in the Twitter discussion.  Which are the top 10 most tagged accounts in the data set. Hint: the "explore_hashtags" chunk is a good starting point.
```{r create_corpus}
corpus <- corpus(dat$Title) #enter quanteda
# corpus is a collection of documents (ie tweets) with metadata
```

```{r explore_hashtags}
tagged_tweets <- tokens(corpus, remove_punct = TRUE) %>% 
               tokens_keep(pattern = "@*")
dfm_tagged<- dfm(tagged_tweets)

tstat_freq <- textstat_frequency(dfm_tagged, n = 10)

#tidytext gives us tools to convert to tidy from non-tidy formats
tagged_tib<- tidy(dfm_tagged)

tagged_tib %>%
   count(term) %>%
   with(wordcloud(term, n, max.words = 10))
```
```{r warning=FALSE}
top_ten_tags <- tagged_tib %>% 
  count(term) %>% 
  top_n(10, n) %>% 
  kable(col.names = c("Tag", "Count")) %>% 
  kable_paper(full_width = FALSE) %>% 
  row_spec(c(0), background = "lightgray")
top_ten_tags
```

# 5. The Twitter data download comes with a variable called "Sentiment" that must be calculated by Brandwatch.  Use your own method to assign each tweet a polarity score (Positive, Negative, Neutral) and compare your classification to Brandwatch's (hint: you'll need to revisit the "raw_tweets" data frame). 
```{r}
dat2<- raw_tweets[,c(4, 6, 10)] # Extract Date, Title, and Sentiment fields

tweets2 <- tibble(text = dat2$Title,
                  element_id = seq(1:length(dat2$Title)),
                 date = as.Date(dat2$Date,'%m/%d/%y'),
                 sent_brandwatch = dat2$Sentiment)
```

```{r warning=FALSE}
sent_method2 <- sentiment_by(tweets2$text)

tweets2 <- inner_join(tweets2, sent_method2, by = "element_id") %>%
  mutate(sent_method2 = case_when(
    ave_sentiment < 0 ~ "negative",
    ave_sentiment > 0 ~ "positive",
    ave_sentiment == 0 ~ "neutral"))
```

```{r}
sent_method_comparison <- tweets2 %>% 
  mutate(sent_comparison = case_when(
    sent_brandwatch == "positive" & sent_method2 == "positive" ~ "both positive",
    sent_brandwatch == "negative" & sent_method2 == "negative" ~ "both negative",
    sent_brandwatch == "neutral" & sent_method2 == "neutral" ~ "both neutral",
    sent_brandwatch == "positive" & sent_method2 == "negative" ~ "brandwatch positive, method 2 negative",
    sent_brandwatch == "positive" & sent_method2 == "neutral" ~ "brandwatch positive, method 2 neutral",
    sent_brandwatch == "neutral" & sent_method2 == "positive" ~ "brandwatch neutral, method 2 positive",
    sent_brandwatch == "neutral" & sent_method2 == "negative" ~ "brandwatch neutral, method 2 negative",
    sent_brandwatch == "negative" & sent_method2 == "positive" ~ "brandwatch negative, method 2 positive",
    sent_brandwatch == "negative" & sent_method2 == "neutral" ~ "brandwatch negative, method 2 neutral"))

sent_method_comparison_counts <- sent_method_comparison %>% 
  count(sent_comparison)
```

```{r warning=FALSE}
sent_method_comparison_counts2 <- sent_method_comparison %>%
  group_by(sent_brandwatch, sent_method2) %>% 
  summarise(count = n())
```


```{r warning=FALSE}
n_tweets <- nrow(tweets2)
sent_method_comparison_table <- sent_method_comparison %>% 
  count(sent_comparison) %>% 
  mutate(percent = round((n / n_tweets) * 100, 1)) %>% 
  kable(col.names = c("Comparison", "Count", "Percent")) %>% 
  kable_paper(full_width = FALSE) %>% 
  row_spec(c(0), background = "lightgray")
sent_method_comparison_table
```
```{r}
both_neg <- sent_method_comparison_counts$n[sent_method_comparison_counts$sent_comparison == "both negative"]
both_pos <- sent_method_comparison_counts$n[sent_method_comparison_counts$sent_comparison == "both positive"]
both_neutral <- sent_method_comparison_counts$n[sent_method_comparison_counts$sent_comparison == "both neutral"]
bw_neu_meth2_pos <- sent_method_comparison_counts$n[sent_method_comparison_counts$sent_comparison == "brandwatch neutral, method 2 positive"]
bw_neu_meth2_neg <- sent_method_comparison_counts$n[sent_method_comparison_counts$sent_comparison == "brandwatch neutral, method 2 negative"]
opposite <- (sent_method_comparison_counts$n[sent_method_comparison_counts$sent_comparison == "brandwatch negative, method 2 positive"]) + (sent_method_comparison_counts$n[sent_method_comparison_counts$sent_comparison == "brandwatch positive, method 2 negative"])
```

There were `r both_neg` tweets where both methods assigned a negative sentiment, `r both_pos` tweets where both methods assigned a positive sentiment, and `r both_neutral` tweets where both methods assigned a neutral sentiment. The greatest disagreements were when brandwatch assigned a neutral sentiment but the other method assigned a positive sentiment (`r bw_neu_meth2_pos` tweets) or a negative sentiment (`r bw_neu_meth2_neg` tweets). There were `r opposite` tweets where the two methods assigned completely opposite sentiments.

```{r}
ggplot(data = sent_method_comparison_counts2, aes(x = sent_brandwatch, y = sent_method2)) +
  geom_tile(aes(fill = count), show.legend = FALSE) +
  geom_text(aes(label = count), color = "black", size = 8) + 
  theme_minimal() +
  theme(panel.grid.major = element_blank()) +
  scale_fill_gradientn(colors = c("seagreen1", "seagreen4")) + 
  labs(title = "Comparison of Sentiment Methods",
       x = "brandwatch sentiment",
       y = "other sentiment method")
  
```