---
title: "assignment_6_word_embeddings"
author: "Marie Rivers"
date: '2022-05-11'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include = FALSE}
library(here)
library(tidytext)
library(tidyverse)
library(widyr)
library(irlba) #singluar value decomposition
library(broom) # creating search_synonym function
library(textdata)
library(ggplot2)
library(dplyr)
library(data.table)
library(kableExtra)
library(patchwork)
```

Read in data
Download a set of pretrained vectors, GloVe, and explore them. 

Grab data here:
<!-- download.file('<https://nlp.stanford.edu/data/glove.6B.zip>',destfile = 'glove.6B.zip')  -->
<!-- unzip('glove.6B.zip')  -->
<!-- Use this file: 'glove.6B.300d.txt' -->

```{r data, message=FALSE, warning=FALSE}
incidents_df<-read_csv("https://raw.githubusercontent.com/MaRo406/EDS_231-text-sentiment/825b159b6da4c7040ce8295b9eae2fbbe9991ffd/dat/climbing_report_text.csv")
```

```{r message=FALSE, warning=FALSE}
glove_data <- fread(here("data", "glove.6B.300d.txt"), header = FALSE) 
glove_df <- glove_data %>%
     remove_rownames() %>%
     column_to_rownames(var = 'V1')
```

First, let's calculate the unigram probabilities, how often we see each word in this corpus.

```{r unigrams}
unigram_probs <- incidents_df %>%
    unnest_tokens(word, Text) %>%
    anti_join(stop_words, by = 'word') %>%
    count(word, sort = TRUE) %>%
    mutate(p = n / sum(n)) 
unigram_probs 
```

Next, we need to know how often we find each word near each other word -- the skipgram probabilities. This is where we use the sliding window.

```{r}
skipgrams <- incidents_df %>%
    unnest_tokens(ngram, Text, token = "ngrams", n = 5) %>%
    mutate(ngramID = row_number()) %>% 
    tidyr::unite(skipgramID, ID, ngramID) %>%
    unnest_tokens(word, ngram) %>%
    anti_join(stop_words, by = 'word')
skipgrams
```

```{r}
#calculate probabilities
skipgram_probs <- skipgrams %>%
    pairwise_count(word, skipgramID, diag = TRUE, sort = TRUE) %>%
    mutate(p = n / sum(n))
```

Having all the skipgram windows lets us calculate how often words together occur within a window, relative to their total occurrences in the data. We do this using the point-wise mutual information (PMI). It's the logarithm of the probability of finding two words together, normalized for the probability of finding each of the words alone. PMI tells us which words occur together more often than expected based on how often they occurred on their own.

```{r norm-prob}
#normalize probabilities
normalized_prob <- skipgram_probs %>%
    filter(n > 20) %>%
    rename(word1 = item1, word2 = item2) %>%
    left_join(unigram_probs %>%
                  select(word1 = word, p1 = p),
              by = "word1") %>%
    left_join(unigram_probs %>%
                  select(word2 = word, p2 = p),
              by = "word2") %>%
    mutate(p_together = p / p1 / p2)
#Which words are most associated with "rope"?   
normalized_prob %>% 
    filter(word1 == "rope") %>%
    arrange(-p_together)
```

Now we convert to a matrix so we can use matrix factorization and reduce the dimensionality of the data.
```{r pmi}
pmi_matrix <- normalized_prob %>%
    mutate(pmi = log10(p_together)) %>%
    cast_sparse(word1, word2, pmi)    
 
#remove missing data
pmi_matrix@x[is.na(pmi_matrix@x)] <- 0
#run SVD using irlba() which is good for sparse matrices
pmi_svd <- irlba(pmi_matrix, 100, maxit = 500) #Reducing to 100 dimensions
#next we output the word vectors:
word_vectors <- pmi_svd$u
rownames(word_vectors) <- rownames(pmi_matrix)
```

1.  Recreate the analyses in the last three chunks (find-synonyms, plot-synonyms, word-math) with the GloVe embeddings. How are they different from the embeddings created from the climbing accident data? Why do you think they are different?

```{r syn-function}
search_synonyms <- function(word_vectors, selected_vector) {
dat <- word_vectors %*% selected_vector
# %*% is matrix multiplication
    
similarities <- dat %>%
        tibble(token = rownames(dat), similarity = dat[,1])
similarities %>%
       arrange(-similarity) %>%
        select(c(2,3))
}
```

```{r find-synonyms-glove}
glove_matrix <- as.matrix(glove_df)

fall_glove <- search_synonyms(glove_matrix, glove_matrix["fall",]) %>% 
  rename(glove_token = token) %>% 
  rename(glove_similarity = similarity) %>% 
  head(15)
slip_glove <- search_synonyms(glove_matrix, glove_matrix["slip",]) %>% 
  rename(glove_token = token) %>% 
  rename(glove_similarity = similarity) %>% 
  head(15)
```

```{r find-synonyms-climb}
fall_climb <- search_synonyms(word_vectors,word_vectors["fall",]) %>% 
  rename(climb_token = token) %>% 
  rename(climb_similarity = similarity) %>% 
  head(15)
slip_climb <- search_synonyms(word_vectors,word_vectors["slip",]) %>% 
  rename(climb_token = token) %>% 
  rename(climb_similarity = similarity) %>% 
  head(15)
```

```{r}
fall_synonyms <- cbind(fall_glove, fall_climb) %>% 
  kable(col.names = c("glove token", "glove similarity", "climb token", "climb similarity"), caption = "Fall Synonyms")
fall_synonyms
```

```{r}
slip_synonyms <- cbind(slip_glove, slip_climb) %>% 
  kable(col.names = c("glove token", "glove similarity", "climb token", "climb similarity"), caption = "Slip Synonyms")
slip_synonyms
```
The glove generated synonyms for fall and slip seem more like true synonyms than those generated from the climbing incident data. This is likely due to the fact that the glove dataset is much larger and includes many more unique words. The climb synonyms for fall and slip seem to be words associated with the lead up and aftermath of the fall or slip (ie line, ice, injuries, fatal). All the top synonyms from the glove data have larger similarity scores than the climbing incident data.

```{r plot-synonyms, fig.height=7, fig.width=8}
glove_synonym_plot <- slip_glove %>%
    mutate(selected = "slip") %>%
    bind_rows(fall_glove %>%
                  mutate(selected = "fall")) %>%
    mutate(glove_token = reorder(glove_token, glove_similarity)) %>%
    ggplot(aes(glove_token, glove_similarity, fill = selected)) +
    geom_col(show.legend = FALSE) +
    facet_wrap(~selected, scales = "free") +
    coord_flip() +
    theme(strip.text=element_text(hjust=0, size=12)) +
    scale_y_continuous(expand = c(0,0)) +
    labs(x = NULL, title = "GloVe word vectors most similar to slip or fall")

climb_synonym_plot <- slip_climb %>%
    mutate(selected = "slip") %>%
    bind_rows(fall_climb %>%
                  mutate(selected = "fall")) %>%
    mutate(climb_token = reorder(climb_token, climb_similarity)) %>%
    ggplot(aes(climb_token, climb_similarity, fill = selected)) +
    geom_col(show.legend = FALSE) +
    facet_wrap(~selected, scales = "free") +
    coord_flip() +
    theme(strip.text=element_text(hjust=0, size=12)) +
    scale_y_continuous(expand = c(0,0)) +
    labs(x = NULL, title = "Climbing incident word vectors most similar to slip or fall")

synonym_plot <- glove_synonym_plot / climb_synonym_plot
synonym_plot
```

```{r word-math-climb}
snow_danger_climb <- word_vectors["snow",] + word_vectors["danger",] 
search_synonyms(word_vectors, snow_danger_climb)
no_snow_danger_climb <- word_vectors["danger",] - word_vectors["snow",] 
search_synonyms(word_vectors, no_snow_danger_climb)
```

```{r word-math-glove}
snow_danger_glove <- glove_matrix["snow",] + glove_matrix["danger",] 
search_synonyms(glove_matrix, snow_danger_glove)
no_snow_danger_glove <- glove_matrix["danger",] - glove_matrix["snow",] 
search_synonyms(glove_matrix, no_snow_danger_glove)
```


2.  Run the classic word math equation, "king" - "man" = ?
```{r king-man}
king_man_glove <- glove_matrix["king",] - glove_matrix["man",] 
search_synonyms(glove_matrix, king_man_glove) %>% 
  head(15)
```

3.  Think of three new word math equations. They can involve any words you'd like, whatever catches your interest.

```{r}
lake_fish <- glove_matrix["lake",] + glove_matrix["fish",] 
search_synonyms(glove_matrix, lake_fish) %>% 
  head(15)
```

```{r}
surf_no_ocean <- glove_matrix["surf",] - glove_matrix["ocean",] 
search_synonyms(glove_matrix, surf_no_ocean) %>% 
  head(15)
```

```{r}
house_no_roof <- glove_matrix["house",] - glove_matrix["roof",] 
search_synonyms(glove_matrix, house_no_roof) %>% 
  head(15)
```

