---
title: 'Topic 6: Topic Analysis'
author: "Marie Rivers"
date: '2022-05-04'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(pdftools)
library(quanteda)
library(tm)
library(topicmodels)
library(ldatuning)
library(tidyverse)
library(tidytext)
library(reshape2)
```

Load the data
```{r}
##Topic 6 .Rmd here:https://raw.githubusercontent.com/MaRo406/EDS_231-text-sentiment/main/topic_6.Rmd
#grab data here: 
comments_df<-read_csv("https://raw.githubusercontent.com/MaRo406/EDS_231-text-sentiment/main/dat/comments_df.csv")

#comments_df <- read_csv(here("dat", "comments_df.csv")) #if reading from local
```
 
Now we’ll build and clean the corpus
```{r}
epa_corp <- corpus(x = comments_df, text_field = "text")
epa_corp.stats <- summary(epa_corp)
head(epa_corp.stats, n = 25)
```

```{r}
toks <- tokens(epa_corp, remove_punct = TRUE, remove_numbers = TRUE)
#I added some project-specific stop words here
add_stops <- c(stopwords("en"),"environmental", "justice", "ej", "epa", "public", "comment")
toks1 <- tokens_select(toks, pattern = add_stops, selection = "remove")
```

And now convert to a document-feature matrix
```{r}
dfm_comm<- dfm(toks1, tolower = TRUE)
dfm <- dfm_wordstem(dfm_comm)
dfm <- dfm_trim(dfm, min_docfreq = 2) #remove terms only appearing in one doc (min_termfreq = 10)

print(head(dfm))
```

```{r}
#remove rows (docs) with all zeros...for the topic model you can't have zeros
sel_idx <- slam::row_sums(dfm) > 0 
dfm <- dfm[sel_idx, ]
#comments_df <- dfm[sel_idx, ]
```

We somehow have to come up with a value for k,the number of latent topics present in the data. How do we do this? There are multiple methods. Let’s use what we already know about the data to inform a prediction. The EPA has 9 priority areas: Rulemaking, Permitting, Compliance and Enforcement, Science, States and Local Governments, Federal Agencies, Community-based Work, Tribes and Indigenous People, National Measures. Maybe the comments correspond to those areas?

```{r}
k <- 9 # k is the number of topics

topicModel_k9 <- LDA(dfm, k, method="Gibbs", control=list(iter = 500, verbose = 25))
```

```{r}
#nTerms(dfm_comm) 

tmResult <- posterior(topicModel_k9)
attributes(tmResult)
```

```{r}
#nTerms(dfm_comm)   
beta <- tmResult$terms   # get beta from results
dim(beta)                # K distributions over nTerms(DTM) terms# lengthOfVocab
```

```{r}
terms(topicModel_k9, 10)
```

Some of those topics seem related to the cross-cutting and additional topics identified in the EPA’s response to the public comments:

1. Title VI of the Civil Rights Act of 1964

2.EJSCREEN

3. climate change, climate adaptation and promoting greenhouse gas reductions co-benefits

4. overburdened communities and other stakeholders to meaningfully, effectively, and transparently participate in aspects of EJ 2020, as well as other agency processes

5. utilize multiple Federal Advisory Committees to better obtain outside environmental justice perspectives

6. environmental justice and area-specific training to EPA staff

7. air quality issues in overburdened communities

So we could guess that there might be a 16 topics (9 priority + 7 additional). Or we could calculate some metrics from the data.
```{r}
result <- FindTopicsNumber(
  dfm,
  topics = seq(from = 2, to = 20, by = 1),
  metrics = c("CaoJuan2009",  "Deveaud2014"),
  method = "Gibbs",
  control = list(seed = 77),
  verbose = TRUE
)

FindTopicsNumber_plot(result)
```
Ran 2 metrics (metrics are different measures of how good the model is)
For the caojuan metric, 20 topics is best
For Deveaud metric, the best number of topics is 7or 10 or 14...

```{r}
k <- 7

topicModel_k7 <- LDA(dfm, k, method="Gibbs", control=list(iter = 500, verbose = 25))
```

```{r}
tmResult <- posterior(topicModel_k7)
terms(topicModel_k7, 10)
```

```{r}
theta <- tmResult$topics
beta <- tmResult$terms
vocab <- (colnames(beta))
```

There are multiple proposed methods for how to measure the best k value. You can go down the rabbit hole here

```{r}
comment_topics <- tidy(topicModel_k7, matrix = "beta")

top_terms <- comment_topics %>%
  group_by(topic) %>%
  top_n(10, beta) %>%
  ungroup() %>%
  arrange(topic, -beta)

top_terms
# beta is the probability of a term in a topic...highest beta or words most likely to be in topic
```

```{r}
top_terms %>%
  mutate(term = reorder(term, beta)) %>%
  ggplot(aes(term, beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  coord_flip()
```

Let’s assign names to the topics so we know what we are working with. We can name them by their top terms

```{r}
top5termsPerTopic <- terms(topicModel_k7, 5)
topicNames <- apply(top5termsPerTopic, 2, paste, collapse=" ")
```

We can explore the theta matrix, which contains the distribution of each topic over each document

```{r}
exampleIds <- c(1, 2, 3)
N <- length(exampleIds)
```

```{r}
#lapply(epa_corp[exampleIds], as.character) #uncomment to view example text
# get topic proportions form example documents
topicProportionExamples <- theta[exampleIds,]
colnames(topicProportionExamples) <- topicNames
vizDataFrame <- melt(cbind(data.frame(topicProportionExamples), document=factor(1:N)), variable.name = "topic", id.vars = "document")  
ggplot(data = vizDataFrame, aes(topic, value, fill = document), ylab = "proportion") +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +  
  coord_flip() +
  facet_wrap(~ document, ncol = N)
# named topics based on first 5 words
# first column show prevalence of each topic in the 1st document
# 1, 2, 3 are the 1st, 2nd, and 3rd documents and the plot shows how each topic is distributed within each document
```

Here’s a neat JSON-based model visualizer

```{r}
library(LDAvis)
library("tsne")
svd_tsne <- function(x) tsne(svd(x)$u)
json <- createJSON(
  phi = tmResult$terms, 
  theta = tmResult$topics, 
  doc.length = rowSums(dfm), 
  vocab = colnames(dfm), 
  term.frequency = colSums(dfm),
  mds.method = svd_tsne,
  plot.opts = list(xlab="", ylab="")
)

serVis(json)
```
each circle is a topic. the size of the circle corresponds to how often the topic appears. If you have a lot of topics, some of the circles might overlap
