---
title: "Untitled"
author: "Marie Rivers"
date: "4/6/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Topic 2: Text Data in R
# Connect to the New York Times API and send a query

```{r}
library(jsonlite) #convert results from API queries into R-friendly formats 
library(tidyverse) 
library(tidytext) #text data management and analysis
library(ggplot2) #plot word frequencies and publication dates
```
q=xxx for query equals
API key

```{r}
#create an object called x with the results of our query ("haaland")
# the from JSON flatten the JSON object, then convert to a data frame
api_key <- # article search api
q <- # query term
t <- fromJSON("http://api.nytimes.com/svc/search/v2/articlesearch.json?q=haaland&api-key=GW3whn8dTpvpdcAD1AIAiUMix03szFDn", flatten = TRUE) #the string following "key=" is your API key 

class(t) #what type of object is t?

t <- t %>% 
  data.frame()
```

```{r}
#Inspect our data
class(t) #now what is it?
dim(t) # how big is it?
names(t) # what variables are we working with?
#t <- readRDS("nytDat.rds") #in case of API emergency :)
```

```{r}
t$response.docs.snippet[9]
# $response refers to the 'response column' from the api results
# snippets (from the NY Times) are sentences. `snippet[9]` pulls the 9th sentence
# try `response.docs.lead_paragraph` to get paragraphs

#assign a snippet to x to use as fodder for stringr functions.  You can follow along using the sentence on the next line.
```


```{r}
x <- "Her nomination as secretary of the interior is historic, but as the first Native cabinet member, she would have to strike a delicate balance."
```


```{r}
tolower(x)
str_split(x, ','); str_split(x, 't')
str_replace(x, 'historic', 'without precedent')
str_replace(x, ' ', '_') #first one
#how do we replace all of them?

str_detect(x, 't'); str_detect(x, 'tive') ### is pattern in the string? T/F
str_locate(x, 't'); str_locate_all(x, 'as')
```

## set some parameters for a bigger query
```{r}
term <- "Haaland" # Need to use + to string together separate words
begin_date <- "20210120"
end_date <- "20220401"

#construct the query url using API operators
baseurl <- paste0("http://api.nytimes.com/svc/search/v2/articlesearch.json?q=",term,
                  "&begin_date=",begin_date,"&end_date=",end_date,
                  "&facet_filter=true&api-key=","NTKBHbsb6XFEkGymGumAiba7n3uBvs8V", sep="")

#examine our query url
```

```{r}
#this code allows for obtaining multiple pages of query results 
 initialQuery <- fromJSON(baseurl)
maxPages <- round((initialQuery$response$meta$hits[1] / 10)-1) # might time out at 8 or 9 pages

pages <- list()
for(i in 0:maxPages){
  nytSearch <- fromJSON(paste0(baseurl, "&page=", i), flatten = TRUE) %>% data.frame() 
  message("Retrieving page ", i)
  pages[[i+1]] <- nytSearch 
  Sys.sleep(6) # change Sys.sleep to 6
}
class(nytSearch)
```

```{r}
# this might be a good place to export to csv as a backup in case the api times out
write_csv(nyData, here("data", "example_nyDat.csv"))
```

```{r}
# xxx
example_nytDat <- read_csv("data/example_nyDat.csv")
```


```{r}
#nytDat <- read.csv("nytDat.csv") # obtained from 

example_nytDat %>% 
  group_by(response.docs.type_of_material) %>%
  summarize(count=n()) %>%
  mutate(percent = (count / sum(count))*100) %>%
  ggplot() +
  geom_bar(aes(y=percent, x=response.docs.type_of_material, fill=response.docs.type_of_material), stat = "identity") + coord_flip()
```

```{r}
example_nytDat %>%
  mutate(pubDay=gsub("T.*","",response.docs.pub_date)) %>%
  group_by(pubDay) %>%
  summarise(count=n()) %>%
  filter(count >= 2) %>%
  ggplot() +
  geom_bar(aes(x=reorder(pubDay, count), y=count), stat="identity") + coord_flip()
```

The New York Times doesn’t make full text of the articles available through the API. But we can use the first paragraph of each article.

The NY Times includes 33 variables (paragraph, author info,...)
We added a 34th column for word as part of the unnesting process (started as 1 row per paragraph, then make a row per word)
```{r}
names(nytDat)
```

```{r}
paragraph <- names(nytDat)[6] #The 6th column, "response.doc.lead_paragraph", is the one we want here.  
tokenized <- nytDat %>%
  unnest_tokens(word, paragraph)
```

```{r}
tokenized %>%
  count(word, sort = TRUE) %>%
  filter(n > 5) %>% #illegible with all the words displayed, consider increasing threshold to 10
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word)) +
  geom_col() +
  labs(y = NULL)
```

## remove stop words
```{r}
data(stop_words)

tokenized <- tokenized %>%
  anti_join(stop_words)

tokenized %>%
  count(word, sort = TRUE) %>%
  filter(n > 5) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word)) +
  geom_col() +
  labs(y = NULL)
```

```{r}
clean_tokens <- str_replace_all(tokenized$word,"land[a-z,A-Z]*","land") #stem tribe words
clean_tokens <- str_remove_all(clean_tokens, "[:digit:]") #remove all numbers
clean_tokens <- str_remove_all(clean_tokens, "washington")
clean_tokens <- gsub("’s", '', clean_tokens)

tokenized$clean <- clean_tokens
```

```{r}
tokenized %>%
  count(clean, sort = TRUE) %>%
  filter(n > 10) %>% #illegible with all the words displayed
  mutate(clean = reorder(clean, n)) %>%
  ggplot(aes(n, clean)) +
  geom_col() +
  labs(y = NULL)
```

```{r}
#remove the empty strings
tib <-subset(tokenized, clean!="")

#reassign
tokenized <- tib

#try again
tokenized %>%
  count(clean, sort = TRUE) %>%
  filter(n > 10) %>% #illegible with all the words displayed
  mutate(clean = reorder(clean, n)) %>%
  ggplot(aes(n, clean)) +
  geom_col() +
  labs(y = NULL)
```


